---
title: "Available sebastes samples"
output: html_notebook
---

Taking a look at data from Ali Deary...

```{r load-data-and-libraries}
library(tidyverse)
library(marmap)
library(hablar)
library(gridExtra)
library(sf)
library(sp)
library(adehabitatHR)

rockfish_df <- readxl::read_xlsx("../data/genetics_rockfish_specimen_list.xlsx", sheet = 2) %>%
  rename(lat = Lat., lon = Long.)

```



```{r}
devtools::install_github("ropensci/rnaturalearthdata")
install.packages("rnaturalearthhires",
                 repos = "http://packages.ropensci.org",
                 type = "source")
```

I think the variables that I want to play with are:
Year
Month
Station
Number

The Max.Gear.Depth and Bottom.Depth could also be interesting, but maybe will be correlated to the Station?

```{r}
rockfish_df %>%
  dplyr::select(Year, Month, Station, Number) %>%
  ggplot() +
  geom_bar(aes(x = Month, y = Number, fill = Station), stat = "identity") +
  facet_grid(cols = vars(Year)) +
  guides(fill = F) +
  theme_bw() +
  ylab("n samples")

```

Maybe there's binning that could happen, both at the temporal level (months 4-5 vs. 7-8, e.g.) and at the spatial level, depending on station locations?
```{r}
rockfish_df %>%
  dplyr::select(Year, Month, Station, Number) %>%
  group_by(Station) %>%
  add_tally(name = "n_samples_per_station")
# unfortunately, there are no more than 2 samples per station

# let's look at the spatial distribution of stations?
```

## Working with spatial data

https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html
```{r}
library("ggplot2")
theme_set(theme_bw())
library("sf")

library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)


ggplot(data = world) +
    geom_sf() +
    coord_sf(crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs ")

```
```{r}
#plot stations lat/lon
plot(rockfish_df$lon, rockfish_df$lat, asp = 1, cex = 0.5, pch = 19)
  
```

```{r}
# antimeridian region 
aleu <- getNOAA.bathy(175, -135, 50, 65, resolution = 4,
antimeridian = TRUE)

# Make it a raster
bathy <- as.raster(aleu)

# Create a xyz table for ggplot
bath<-fortify(aleu)
```


```{r}
library(oce)
library(ocedata)
data("coastlineWorldFine")

# convert bathymetry
bathyLon = as.numeric(rownames(aleu))
bathyLat = as.numeric(colnames(aleu))
bathyZ = as.numeric(aleu)
dim(bathyZ) = dim(aleu)

#filter NAs
rockfish_spatial_df <- rockfish_df %>%
  filter(!is.na(lon))

# define plotting region
mlon = mean(rockfish_spatial_df$lon)
mlat = mean(rockfish_spatial_df$lat)
span = 2000
```


Plot it up
```{r plot}
# set colors to the Station identity
rockfish_spatial_df %>%
  dplyr::select(Station) %>%
  unique()

library(RColorBrewer)
# Define the number of colors you want
nb.cols <- 101
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)
# Create a ggplot with 101 colors
```


```{r plot-station-locs}
# set up the output
pdf("pdf_outputs/station_map.pdf", width = 8, height = 6)

# plot station locations
plot(coastlineWorldFine, clon = mlon, clat = mlat, span = 1100, 
     projection="+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154", col = 'lightgrey')

# plot bathymetry
mapContour(bathyLon, bathyLat, bathyZ,
           levels = c(-500, -1000, -2500),
           lwd = c(1.5, 1, 1),
           lty = c(3, 1, 3),
           col = 'darkgray')

# add depth legend
#legend("topright", seg.len = 3, cex = 0.7,
#        lwd = c(1.5, 1, 1),
#        lty = c(3, 1, 3),
#        legend = c("-500", "-1000", "-2500"),
#        box.col = "white", col = 'darkgray', title = "Depth (m)", bg = "white")

# add map data
# add colony locations
mapPoints(longitude = rockfish_spatial_df$lon, latitude = rockfish_spatial_df$lat, pch = 19, col = mycolors, cex = 0.5, lwd = 2)
#text(longitude = rockfish_spatial_df$lon, latitude = rockfish_spatial_df$lat, rockfish_spatial_df$Station, pos=4)

dev.off()
```

Save that map as an object to add to my grid?
```{r}
# grid libraries
library(gridExtra)
library(gridGraphics)
library(grid)

# plot station locations
plot(coastlineWorldFine, clon = mlon, clat = mlat, span = 1100, 
     projection="+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154", col = 'lightgrey')

# plot bathymetry
mapContour(bathyLon, bathyLat, bathyZ,
           levels = c(-500, -1000, -2500),
           lwd = c(1.5, 1, 1),
           lty = c(3, 1, 3),
           col = 'darkgray')

# add locations
mapPoints(longitude = rockfish_spatial_df$lon, latitude = rockfish_spatial_df$lat, pch = 19, col = mycolors, cex = 0.5, lwd = 2)
# to save as an object
grid.echo()
map.fig <- grid.grab()

# let's see if that works to add into the panel of plots
```


```{r}
rockfish_spatial_df %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, color = Station)) +
  #geom_text(aes(x = lon, y = lat, label = Station)) +
  ggrepel::geom_text_repel(aes(x = lon, y = lat, label = Station)) +
  guides(color = F)

ggsave("pdf_outputs/station_labeled_plot.pdf", width = 10, height = 6)
```

How to create defined clusters from this data?

https://rpubs.com/aephidayatuloh/clustervisual
```{r}
# kmeans method

rock_cluster <- rockfish_spatial_df %>%
  dplyr::select(lat, lon)

cls <- kmeans(x = rock_cluster, centers = 6)
rock_cluster$cluster <- as.character(cls$cluster)
head(rock_cluster)


ggplot() +
  geom_point(data = rock_cluster, 
             mapping = aes(x = lon, 
                                  y = lat, 
                                  colour = cluster))
```

http://www.supplychaindataanalytics.com/proximity-based-spatial-customer-grouping-in-r/

Use K-means clustering, but first define starting points based on the longitude dimensions of the spatial dataset
```{r}
# define a fct
initial_centers <- function(stations,centers){
  quantiles <- c()
  for(i in 1:centers){
    quantiles <- c(quantiles,i*as.integer(nrow(stations)/centers))
  }
  quantiles
}

```

```{r}
cluster_obj <- kmeans(rock_cluster,centers=rock_cluster[initial_centers(rock_cluster,6),])
head(cluster_obj)
```


```{r}
result_df <- rock_cluster
result_df$group <- cluster_obj$cluster
head(result_df)

result_df <- result_df %>%
  convert(., chr(group))

```


```{r}
grp_rockfish_df <- result_df %>%
  left_join(., rockfish_spatial_df, by = c("lat", "lon")) 

station_locs_plot <- grp_rockfish_df %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, color = group)) +
  #geom_text(aes(x = lon, y = lat, label = Station)) +
 ggrepel::geom_text_repel(aes(x = lon, y = lat, label = Station)) +
  #guides(color = F) 
  ylab("Latitude") +
  xlab("Longitude") +
  theme(
    legend.position = "top"
  )

station_locs_plot
```

Okay, so if I use these 6 groups, what kind of temporal coverage do I get?

What about by month?
```{r}
# make month a factor
grp_rockfish_df$Month <- factor(grp_rockfish_df$Month)

# count the data
count_data = count(grp_rockfish_df, Month, group, Year, Number)
# fill in the missing combination
count_data = complete(count_data, Month, group, Year, Number) 

count_data %>%
  arrange(Year)
```

```{r}
# set month names for labeling?
month.labs <- c("April", "May", "June", "August")
names(month.labs) <- c("4","5","6","8")

# plot number of samples by station group, month, and year
samples_mo_yr_station_facet <- grp_rockfish_df %>%
  dplyr::select(Month, group, Year, Number) %>%
  count(Month, group, Year, Number) %>%
  complete(Month, Year, group) %>%
  ggplot(aes(x = group, y = Number, fill = group)) +
  geom_bar(stat = "identity", position = "dodge") +
  #facet_grid(rows = vars(Year), cols = vars(Month), labeller = labeller(.cols = label_both)) +
  facet_grid(rows = vars(Year), cols = vars(Month), labeller = labeller(Month = month.labs), scales = "free_y") +
  guides(fill = F) +
  xlab("Station group") +
  ylab("Number of samples") #+
  #theme(
   # legend.position = "bottom"
  #)

#ggsave("pdf_outputs/rockfish_samples_by_month_year_grp.pdf", height = 6, width = 7)
```

Now, the thing to do is combine this panel with the map so that the station group locations are evident.

https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html

```{r multi-plot-page}

panel <- grid.arrange(samples_mo_yr_station_facet, station_locs_plot, nrow = 2)

ggsave(panel, file = "pdf_outputs/station_grps_panel.pdf", height = 10, width = 6)
```

Now use the map in addition?

```{r multi-plot-page2}
# add the map as well
panel3 <- grid.arrange(samples_mo_yr_station_facet, station_locs_plot, map.fig, nrow = 3)

ggsave(panel3, file = "pdf_outputs/station_grps_panel3.pdf", height = 10, width = 6)
```

This can be refined by:
1. modifying the colors/groups to display in the map
2. adding labels to the map rather than using the plot?


## Spring vs. late summer

Seasonal changes was one question that Ali, Wes, and I discussed looking at with these data.
So the question would be, how many samples fall into those categories by year?

```{r}
season_df <- rockfish_df %>%
  mutate(season = ifelse(Month %in% c(4,5,6), "spring", "late_summer")) %>%
  group_by(season, Year) %>%
  summarise(samples = sum(Number)) 

season_df$season <- factor(season_df$season, levels = c("spring", "late_summer"))
season_df$Year <- factor(season_df$Year)

season_df %>%
  count(season, Year, samples) %>%
  complete(season, Year) %>% # another counting issue to keep the x-axis consistent
  ggplot() +
  geom_bar(aes(x = Year, y = samples, fill = season), stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Dark2")

```


What about treating April differently from June?
```{r}
rockfish_df$Year <- factor(rockfish_df$Year)
rockfish_df$Month <- factor(rockfish_df$Month)

rockfish_df %>%
  group_by(Year, Month) %>%
  summarise(n_samples = sum(Number)) %>%
  # count the missing entries
  count(Year, Month, n_samples) %>%
  complete(Year, Month) %>%
  ggplot() +
  geom_bar(aes(x = Month, y = n_samples, fill = Month), stat = "identity", position = "dodge") +
  facet_grid(rows = vars(Year), scales = "free_y") +
  theme(
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "Dark2")
  

ggsave("pdf_outputs/samples_by_year_month.pdf", width = )
```


If we have ~1000 samples to genotype initially, there are a couple of things I would do:
1. look at composition over the longest time period available (April-August)
2. look at changes in composition across years using best available comparative data (May/June)
3. ignore spatial distribution (for now?)

Assume ~100 samples per Month/Year... how many samples in each category?

April 2013 (927)*
May 2011 (759)*
June 2011 (28)
July 2012 (1277)*
August 2012 (99)

May 2009 (26)
June 2009 (7)
June 2011 (28)
May 2015 (1022)*
June 2015 (645)*

*downsample these

```{r}
rockfish_df %>%
  group_by(Year, Month) %>%
  summarise(n_samples = sum(Number)) %>%
  
```

## Spatial distribution of samples

Based on Goldstein et al., 2019, I could look at the Eastern and Western Gulf of Alaska, with a biogeographic break at -146 degrees (Prince William Sound).

```{r}
# how many samples are available for the different regions for the time periods of interest, above?
goa_region <- rockfish_spatial_df %>%
  mutate(spatial_area = ifelse(lon < -146, "EGOA", "WGOA")) 

goa_region$Year <- factor(goa_region$Year)
goa_region$Month <- factor(goa_region$Month)
goa_region$spatial_area <- factor(goa_region$spatial_area)

# plot that up
region_sample_ct <- goa_region %>%
  group_by(Year, Month, spatial_area) %>%
  summarise(n_samples = sum(Number)) %>%
  # count the missing entries
  count(Year, Month, n_samples, spatial_area) %>%
  complete(Year, Month, spatial_area) %>%
  ggplot() +
  geom_bar(aes(x = spatial_area, y = n_samples, fill = spatial_area), stat = "identity", position = "dodge") +
  facet_grid(rows = vars(Year), cols = vars(Month), labeller = labeller(Month = month.labs), scales = "free_y") +
  theme(
    legend.position = "bottom"
  ) +
  scale_fill_manual(values = c("dark blue", "goldenrod"))
  
ggsave("pdf_outputs/EastWest_regions_sample_ct.pdf", width = 5, height = 6)
```

Can I get that info onto the map?

```{r plot-GoA-regions}
# organize the color palette
region_cols <- c("goldenrod", "dark blue")[goa_region$spatial_area]
  
# plot station locations
plot(coastlineWorldFine, clon = mlon, clat = mlat, span = 1100, 
     projection="+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154", col = 'lightgrey')

# plot bathymetry
mapContour(bathyLon, bathyLat, bathyZ,
           levels = c(-500, -1000, -2500),
           lwd = c(1.5, 1, 1),
           lty = c(3, 1, 3),
           col = 'darkgray')

# add GoA locations
mapPoints(longitude = goa_region$lon, latitude = goa_region$lat, pch = 19, col = region_cols, cex = 0.5, lwd = 2)
title(xlab = "Longitude", ylab = "Latitude")

# to save that output as an object
grid.echo()
map.fig_regions <- grid.grab()

```

And put those two things together as a panel:
```{r panel-the-regions}
# add the map and barplot
GoA_regions_panel <- grid.arrange(map.fig_regions, region_sample_ct, nrow = 1)

ggsave(GoA_regions_panel, file = "pdf_outputs/GoA_regions_panel.pdf", height = 5, width = 11)

```






## Messing around with mapping

convert to a projection that is more appropriate for AK
```{r reproject-datapoints}
coast_tr <- st_transform(structure(coast, proj4string = "+proj=longlat +datum=WGS84 +no_defs"), crs = "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154")


```




```{r}
ggplot(data = coast_tr) +
    geom_sf() +
    coord_sf()
    #coord_sf(xlim = c(-180,-130), ylim = c(45,70), expand = T)

```

```{r}
ne_coastline()

lakes110 <- ne_download(scale = 110, type = 'lakes', category = 'physical')

coasts <- ne_download(scale = 110, type = 'coastline', category = 'physical')
```


```{r}
# what's the best map for my area?
coastlineBest(lonRange = c(-180,-130), latRange = c(45,70), debug = 1)
```

```{r}

?coastlineWorldMedium

data(coastlineWorldMedium, package="ocedata")

sp::plot(coastlineWorldMedium)
sp::plot(ne_coastline())
# coastline of the world
# subsetting of coastline is not possible because the Natural Earth data are not attributed in that way
```

```{r}
st_crs(ne_coastline())

str(ne_coastline)

coasts <- ne_coastline(scale = 100, type = " ", returnclass = "sf")

ggplot(data = coasts) +
  geom_sf(fill = "grey", colour = "black") +
  coord_sf()
```

```{r}
library(rnaturalearth)
library(sf)
library(ggplot2) # with support for geom_sf

# all countries at scale 10m
ctrys <- ne_countries(scale = 10, type = "countries", returnclass = "sf")

crsLONGLAT <- "+proj=longlat +datum=WGS84 +no_defs"
crsLAEA <- "+proj=laea +lat_0=55 +lon_0=-170 +x_0=4321000 +y_0=3210000 +datum=WGS84 +units=m +no_defs"

# the bouding box polygon in long/lat projection, i.e. axis-aligned
bb <- st_sfc(
  st_polygon(list(cbind(
    c(-42, 52, 52, -42, -42), # x-coordinates (longitudes) of points A,B,C,D
    c(12, 12, 84, 84, 12)     # y-coordinates (latitudes) of points A,B,C,D
    ))),
  crs = crsLONGLAT)


# now in in LAEA projection
laeabb <- st_transform(bb, crs = crsLAEA)

# the extent of the bounding box in the new projection
b <- st_bbox(laeabb)
b

gg2 <- ggplot(data = coasts) +
  geom_sf(fill = "grey", colour = "black") +
  coord_sf()


  coord_sf(crs = crsLAEA, xlim = c(b["xmin"], b["xmax"]), ylim = c(b["ymin"], b["ymax"]))
gg2
```

